---
title: "Tidycensus Markdown"
author: "Akira Di Sandro"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE, cache = TRUE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(knitr)
library(kableExtra)
library(rmarkdown)
library(tidycensus)
library(sf)
```

# Census API Key

```{r load_key, warning = FALSE, eval = FALSE}
census_api_key("f2855a6037284cb9cbed55e96e6b99be17ee05c6", overwrite = TRUE)
```

## Load census data dictionaries

```{r load_variables, cache = TRUE}

acs_variable_list.2020 <- load_variables(2020, #year
                                         "acs5", #five year ACS estimates
                                         cache = TRUE)

acs_variable_list.2016 <- load_variables(2016, #year
                                         "acs5", #five year ACS estimates
                                         cache = TRUE)

# use things like glimpse(), str(), summary(), table(dataframe$variable)
```

# Downloading Data from Tidycensus

```{r acs_vars}
acs_vars <- c("B01001_001E", # ACS total Pop estimate
              "B25002_001E", # Estimate of total housing units
              "B25002_003E", # Number of vacant housing units
              "B19013_001E", # Median HH Income ($)
              "B02001_002E", # People describing themselves as "white alone"
              "B06009_006E") # Total graduate or professional degree
```

## Call the Census API to get tract level data for 2020 for all of Philadelphia

```{r get_acs_2020, cache = TRUE, message = FALSE, warning = FALSE}
acsTractsPHL.2020 <- get_acs(geography = "tract",
                             year = 2020, 
                             variables = acs_vars, 
                             geometry = FALSE, 
                             state = "PA", 
                             county = "Philadelphia", 
                             output = "wide") 
```

# Wrangling Data with dplyr

## Mutating, selecting and renaming variables

```{r do_some_dplyr, cache = TRUE}
acsTractsPHL.2020 <- acsTractsPHL.2020 %>%
  dplyr::select (GEOID, NAME, all_of(acs_vars))

acsTractsPHL.2020 <- acsTractsPHL.2020 %>%
  rename (total_pop.2020 = B01001_001E,
          total_HU.2020 = B25002_001E,
          total_vacant.2020 = B25002_003E,
          med_HH_Income.2020 = B19013_001E,
          total_White.2020 = B02001_002E,
          total_GradDeg.2020 = B06009_006E)

acsTractsPHL.2020 <- acsTractsPHL.2020 %>%
  mutate(vacancyPct.2020 = total_vacant.2020/total_HU.2020,
         pctWhite.2020   = total_White.2020/total_pop.2020)
```

```{r get_acs_2016, cache = TRUE, message = FALSE}
acsTractsPHL.2016 <- get_acs(geography = "tract",
                             year = 2016, 
                             variables = acs_vars,
                             geometry = FALSE,
                             state = "PA", 
                             county = "Philadelphia",
                             output = "wide") %>%
  dplyr::select (GEOID, NAME, all_of(acs_vars)) %>% 
  rename (total_pop.2016 = B01001_001E,
          total_HU.2016 = B25002_001E,
          total_vacant.2016 = B25002_003E,
          med_HH_Income.2016 = B19013_001E,
          total_White.2016 = B02001_002E,
          total_GradDeg.2016 = B06009_006E) %>%
  mutate(vacancyPct.2016 = total_vacant.2016/total_HU.2016,
         pctWhite.2016 = total_White.2016/total_pop.2016)
```

## Joining data

```{r left_join_tracts, cache = TRUE}
allACS <- left_join(acsTractsPHL.2016, acsTractsPHL.2020,
                    by= c("GEOID"))
```

## Doing column math using mutate

```{r do_mutates, cache = TRUE}
allACS <- allACS %>%
  mutate(change_med_HH_Income = med_HH_Income.2020 - (med_HH_Income.2016 * 1.08), 
         change_Grad_Degree_Pct = (total_GradDeg.2020/total_pop.2020)-(total_GradDeg.2016/total_pop.2016))

```


# Comparing geographies

```{r myTracts, cache = TRUE}

myTracts <- c("42101023500", 
              "42101023600", 
              "42101023700", 
              "42101025300", 
              "42101025400",
              "42101025500", 
              "42101025600", 
              "42101038800")

allACS <- allACS %>%
  mutate(mtAiry = ifelse(GEOID %in% myTracts, "MT AIRY", "REST OF PHILADELPHIA"))
```

# Spatial Data and Tidycensus

```{r spatial_tidycensus, message=FALSE, warning=FALSE, cache=TRUE, include=FALSE}
acsTractsPHL.2020.sf <- get_acs(geography = "tract",
                             year = 2020, 
                             variables = acs_vars, 
                             geometry = TRUE, 
                             state = "PA", 
                             county = "Philadelphia", 
                             output = "wide") %>% 
  dplyr::select (GEOID, NAME, all_of(acs_vars)) %>%
  rename (total_pop.2020 = B01001_001E,
          total_HU.2020 = B25002_001E,
          total_vacant.2020 = B25002_003E,
          med_HH_Income.2020 = B19013_001E,
          total_White.2020 = B02001_002E,
          total_GradDeg.2020 = B06009_006E) %>%
  mutate(vacancyPct.2020 = total_vacant.2020/total_HU.2020,
         pctWhite.2020 = total_White.2020/total_pop.2020) %>%
  mutate(mtAiry = ifelse(GEOID %in% myTracts, "MT AIRY", "REST OF PHILADELPHIA")) %>%
  st_as_sf(crs = 4326) # Turn shp into sf object and project as WGS84
# 4326 stands for "projection" with lattitude and longitude
```


# Assignment

a) a ggplot plot city-wide - number of vacant housing units per tract in 2020 as a function of vacant housing units in 2016,  (change in raw housing units, so 2020-2016, using "B25002_003E", # Number of vacant housing units", which is part of the `acs_vars` vector). Use the ggplot "color" parameter in the geom_point aesthetics to visualize which are the Mt. Airy tracts.  Annotate this plot with a single sentence describing what it shows (not as a plot title, but within the markdown).

```{r assignment_a, warning = FALSE, cache = TRUE, message = FALSE}
ggplot(allACS) +
  geom_point(aes(x = total_vacant.2016, y = total_vacant.2020,
             color = mtAiry)) +
  labs(
    title = "2020 Vacant Housing units as a Function of 2016 Vacant Housing units",
    caption = "Data: US Census Bureau, ACS 5-year estimates",
    x = "Total Vacant Housing Units in 2016", 
    y = "Total Vacant Housing Units in 2020")
  
```

Plot Annotation:
Though some regions saw drastic changes in total number of vacant housing units between 2016 and 2020, the number of vacant housing units mostly remained consistent in most regions of Philadelphia.

b) A ggplot map symbolizing the number of vacant housing units in 2016 as a percentage of total housing units (e.g.  "total_HU.2016") on a per-tract basis. Outline the Mt. Airy neighborhood on the map. Annotate this map with a single sentence describing what it shows or implies about Mt Airy and Philadelphia.
```{r assignment_b1, warning = FALSE, cache = TRUE, message = FALSE, include = F}
# first create a sf data frame for 2016 ACS data
acsTractsPHL.2016.sf <- get_acs(geography = "tract",
                             year = 2016, 
                             variables = acs_vars, 
                             geometry = TRUE, 
                             state = "PA", 
                             county = "Philadelphia", 
                             output = "wide") %>% 
  dplyr::select (GEOID, NAME, all_of(acs_vars)) %>%
  rename (total_pop.2016 = B01001_001E,
          total_HU.2016 = B25002_001E,
          total_vacant.2016 = B25002_003E,
          med_HH_Income.2016 = B19013_001E,
          total_White.2016 = B02001_002E,
          total_GradDeg.2016 = B06009_006E) %>%
  mutate(vacancyPct.2016 = total_vacant.2016/total_HU.2016,
         pctWhite.2016 = total_White.2016/total_pop.2016) %>%
  mutate(mtAiry = ifelse(GEOID %in% myTracts, "MT AIRY", "REST OF PHILADELPHIA")) %>%
  st_as_sf(crs = 4326)
  
```

```{r assignment_b2, warning = FALSE, cache = TRUE, message = FALSE}
# now create ggplot map
ggplot()+
  geom_sf(data = acsTractsPHL.2016.sf, aes(fill = vacancyPct.2016),
          color = "transparent")+
  geom_sf(data = acsTractsPHL.2016.sf %>%
            filter(mtAiry == "MT AIRY") %>%
            st_union(),
          color = "white",
          fill = "transparent")+
  labs(
    title = "Percentage of Vacant Housing Units in 2016",
    subtitle = "",
    caption = "Data: US Census Bureau, ACS 5-year estimates",
    fill = "% Vacant Units")
  
```

Plot Annotation:
About 20% of the total housing units in the Mt. Airy region are vacant, which can be said about the general Philadelphia area, but there is a higher percentage (closer to 40%) of vacant housing units in the upper West and North Philadelphia regions. 

c) Create a code chunk that renders a "kable" table showing a summary of mean number of vacant housing units per tract for Mt. Airy vs. the rest Philadelphia as a whole in 2020.  Annotate this plot with a single sentence describing what it shows or implies about the data. 

```{r assignment_c, warning = FALSE, cache = TRUE, message = FALSE}
# make desired table
allACS %>%
  group_by(mtAiry) %>%
  summarize(mean_VacantUnits = mean(total_vacant.2020, na.rm = T)) %>% 
  kbl(caption = "Mean Vacant Units per tract in Mt. Airy vs Rest of Philadelphia",
      col.names = c("Region", "Mean number of Vacant Units")) %>%
  kable_classic(full_width = F, html_font = "Cambria")

```

Table Annotation:
This Table shows us that the mean number of vacant units in Mt. Airy is lower compared to that of the rest of Philadelphia.




